### 2014-07-09 text/Chapter04.tex
Finish cbn section, notes on normal forms (Chapter 4)

@@

\subsection{Syntactic properties}

@@

The (\emph{internalized substitution}) rule models an operational semantics where the language
provides a construct that models \emph{sequential composition} and expressions can be reduced to 
a normal form, consisting of a sequenced context-dependent operations. We choose to write the 
sequencing operation as \kvd{glet} to emphasize that this is a separate primitive and not an 
ordinary syntactic \kvd{let}. The normal form looks as follows:
%
\begin{equation*}
\kvd{glet}~x_1 = e_1~\kvd{in}~\ldots~\kvd{glet}~x_n = e_n~\kvd{in}~e
\end{equation*}
%
Here, the expressions $e_1, \ldots, e_n, e$ do not contain further nested \kvd{glet} constructs.
This evaluation strategy provides context-dependent counterpart to operational view of monads
developed by Filinsky \cite{monads-inaction}. We discuss how expressions reduce to the normal form
in Section~\ref{sec:flat-syntax-norm}

The (\emph{internalized substitution}) rule is useful when defining a concrete semantics for 
a language that provides constructs for explicitly providing the context. For example, consider
a language with implicit parameters where a parameter is defined by $e_1~\kvd{with}~\ident{?p}=e_2$.
Semantics for such language would provide a reduction rule:
%
\begin{equation*}
(\kvd{glet}~x_1 = \ident{?p}~\kvd{in}~e)~\kvd{with}~\ident{?p}=e_1 ~~\leadsto~~ \kvd{let}~x_1 = e_1~\kvd{in}~e
\end{equation*}
%
Here, the \kvd{glet} construct provides a way of sequentialising the context-requi\-rements so
that they can be discharged by matching constructs providing the required contexts. As discussed
earlier, we focus on the general case and so we discuss when a flat coeffect calculus supports
this form of evaluation (rather than looking at semantics for concrete systems).

@@

\begin{theorem}[Call-by-value reduction]
In a flat coeffect system with the (\emph{approxi\-mation}) property, 
if $\coctx{\Gamma}{\cclrd{r}} \vdash e : \tau$ 
and $e \reduce_{\ident{cbv}} e'$ then $\coctx{\Gamma}{\cclrd{r}} \vdash e' : \tau$.
\end{theorem}
\begin{proof}
Consider the typing derivation for the term $(\lambda x.e)\,v$ before reduction:
\begin{equation*}
\inference
  { \inference
  { \inference
      {\coctx{\Gamma, x:\tau_1}{\cclrd{r}\;\czip\;\cclrd{t}} \vdash e : \tau_2}
      {\coctx{\Gamma}{\cclrd{r}} \vdash \lambda x.e : \tau_1 \xrightarrow{\cclrd{t}} \tau_2 } &
   \coctx{\Gamma}{\cunit} \vdash v : \tau_1 }
  {\coctx{\Gamma}{\cclrd{r} \;\cpar\; (\cunit \,\cseq\, \cclrd{t})} \vdash (\lambda x.e)\,v : \tau_2} }
  {\coctx{\Gamma}{\cclrd{r} \;\cpar\; \cclrd{t}} \vdash (\lambda x.e)\,v : \tau_2}
\end{equation*}
The last step simplifies the coeffect annotation using the fact that $\cunit$ is a unit of $\cseq$.
From Lemma~\ref{thm:flat-subst-cbv} $\subst{e}{x}{v}$ has the same coeffect annotation as $e$.
As $\cclrd{r}\,\czip\,\cclrd{t}\;\cleq\;\cclrd{r}\,\cpar\,\cclrd{t}$, we can apply sub-coeffecting:
%
\begin{equation*}
\tyrule{sub}
 { \coctx{\Gamma, x:\tau_1}{\cclrd{r}\;\czip\;\cclrd{t}} \vdash \subst{e}{x}{v} : \tau_2 }
 { \coctx{\Gamma, x:\tau_1}{\cclrd{r}\;\cpar\;\cclrd{t}} \vdash \subst{e}{x}{v} : \tau_2 } 
\end{equation*}
%
Thus, the reduction preserves type and coeffect annotation (although this may not be the
\emph{only} typing of the original term).
\end{proof}

@@

The call-by-name strategy reduces a term $(\lambda x. e_1)~e_2$ where both sub-expressions 
may have contextual requirements. Assume that $\coctx{\Gamma}{\cclrd{s}} \vdash e_2 : \tau_2$ and
$\coctx{\Gamma_1, x :\tau_1}{\cclrd{r}} \vdash e_1 : \tau_1$.

% TODO
%
% The context requirements of the 
% reduced term $\subst{e_1}{x}{e_2}$ should be the same as the context requirements of 
% $\kvd{let}~x=e_2~\kvd{in}~e_1$. Thus the context requirements of the result in the substitution 
% lemma should be $r \fmax (s \fseq r)$.

We call a flat coeffect algebra \emph{top-pointed} if $\cunit$ is the greatest (top) coeffect scalar
$\C$ and \emph{bottom-pointed} if it is the smallest (bottom) coeffect scalar.
Liveness analysis is an example of top-pointed coeffects
as variables are annotated with $\ident{L}$ and $\ident{D} \leq \ident{L}$.

\begin{lemma}[Top-pointed substitution]
\label{thm:cbn-substitution-top}
In a top-pointed flat coeffect calculus with $(\C, \cseq, \cpar, \cunit, \czero, \cleq)$ 
and the $\cclrd{\wedge}$ operator:
\[
\begin{array}{l}
\coctx{\Gamma}{\cclrd{s}} \vdash e_s : \sigma \;\; \wedge \;\; \coctx{\Gamma_1,  x : \sigma, \Gamma_2}{ \cclrd{r}  } \vdash e_r : \tau\\
\quad \Rightarrow \;\; \coctx{\Gamma_1,\Gamma,\Gamma_2}{ \cclrd{r} } \vdash \subst{e_r}{x}{e_s} : \tau
\end{array}
\]
\vspace{-1.5em}
\end{lemma}
\begin{proof}
Using sub-coeffecting ($\cclrd{s} \,\cleq\, \cunit$) and Lemma~\ref{thm:flat-subst-cbv}.
\end{proof}
%
\noindent
As variables are annotated with the top element $\cunit$, we can substitute a term $e_s$ for any variable and 
use sub-coeffecting to get the original typing (because $\cclrd{s} \,\cleq\, \cunit$).

In a bottom pointed coeffect system, substituting $e$ for $x$ increases the context 
requirements. However, if the system satisfies the strong condition that $\cclrd{\wedge} = \cseq = \cpar$ then
the context requirements arising from the substitution can be associated with the context
$\Gamma$. As a result, substitution does not break soundness as in effect systems. The 
requirement $\cclrd{\wedge} = \cseq = \cpar$ holds for our implicit parameters example 
(all three operators are set union) and allows the following substitution lemma:

\begin{lemma}[Bottom-pointed substitution]
\label{thm:cbn-substitution-bot}
In a bottom-pointed flat coeffect calculus with $(\C, \cseq, \cpar, \cunit, \czero, \cleq)$ 
and the $\cclrd{\wedge}$ operator where $\cclrd{\wedge} = \cseq = \cpar$ is idempotent and commutative:
\[
\begin{array}{l}
\coctx{\Gamma}{\cclrd{s}} \vdash e_s : \sigma \;\; \wedge \;\; \coctx{\Gamma_1,  x : \sigma, \Gamma_2}{ \cclrd{r}  } \vdash e_r : \tau\\
\quad \Rightarrow \;\; \coctx{\Gamma_1,\Gamma,\Gamma_2}{ \cclrd{r} \,\cseq\, \cclrd{s} } \vdash \subst{e_r}{x}{e_s} : \tau
\end{array}
\]
\vspace{-1.5em}
\end{lemma}
\begin{proof}
By induction over $\vdash$, using the idempotent, commutative monoid structure to keep $\cclrd{s}$ with
the free-variable context.
\end{proof}
\noindent
The structural system is precise enough to keep the coeffects associated with a concrete variable. The flat
variant described here is flexible enough to let us always re-associate new context requirements with the 
free-variable context.

The two substitution lemmas show that the call-by-name evaluation strategy can be used for
certain coeffect calculi, including liveness and implicit parameters.
Assuming $\reduce_{\ident{cbn}}$ is the standard call-by-name
reduction, the following theorem holds:

\begin{theorem}[Call-by-name reduction]
\raggedright
In a coeffect system that satisfies the conditions for Lemma~\ref{thm:cbn-substitution-top} or
Lemma~\ref{thm:cbn-substitution-bot}, if $\coctx{\Gamma}{\cclrd{r}} \vdash e : \tau$ 
and $e \rightarrow_{\ident{cbn}} e'$ then $\coctx{\Gamma}{\cclrd{r}} \vdash e' : \tau$.
\end{theorem}
\begin{proof}
A direct consequence of Lemma~\ref{thm:cbn-substitution-top} or Lemma~\ref{thm:cbn-substitution-bot}.
\end{proof}

ALSO: Something on eta expansion

\subsection{Normal form}
\label{sec:flat-syntax-norm}

@@

@@

or

>>>>

%!TEX root = ../main.tex

@@

\subsection{Syntactic properties}
\label{sec:flat-syntax-props}

@@

The (\emph{internalized substitution}) rule captures an operational semantics where the language
provides a construct representing \emph{sequential composition} and expressions can be reduced to 
a normal form, consisting of a sequenced context-dependent operations. We choose to write the 
sequencing operation as \kvd{glet} to emphasize that this is a separate primitive and not an 
ordinary syntactic \kvd{let}. The normal form looks as follows:
%
\begin{equation*}
\kvd{glet}~x_1 = e_1~\kvd{in}~\ldots~\kvd{glet}~x_n = e_n~\kvd{in}~e
\end{equation*}
%
Here, the expressions $e_1, \ldots, e_n, e$ do not contain further nested \kvd{glet} constructs.
This evaluation strategy provides context-dependent counterpart to operational view of monads
developed by Filinsky \cite{monads-inaction}. We discuss how expressions reduce to the normal form
in Section~\ref{sec:flat-syntax-norm}

The (\emph{internalized substitution}) rule is useful when defining a concrete semantics for 
a language that provides constructs for explicitly providing the context. For example, consider
a language with implicit parameters where a parameter is defined by $e_1~\kvd{with}~\ident{?p}=e_2$.
Semantics for such language would provide a reduction rule:
%
\begin{equation*}
(\kvd{glet}~x_1 = \ident{?p}~\kvd{in}~e)~\kvd{with}~\ident{?p}=e_1 ~~\leadsto~~ \kvd{let}~x_1 = e_1~\kvd{in}~e
\end{equation*}
%
Here, the \kvd{glet} construct provides a way of sequentialising the context-requi\-rements so
that they can be discharged by matching constructs providing the required contexts. As discussed
tfearlier, we focus on the general case and so we discuss when a flat coeffect calculus supports
this form of evaluation (rather than looking at semantics for concrete systems).

@@

\begin{theorem}[Call-by-value reduction]
\label{thm:cbv-reduction}
In a flat coeffect system with the (\emph{approxi\-mation}) property, 
if $\coctx{\Gamma}{\cclrd{r}} \vdash e : \tau$ 
and $e \reduce_{\ident{cbv}} e'$ then $\coctx{\Gamma}{\cclrd{r}} \vdash e' : \tau$.
\end{theorem}
\begin{proof}
Consider the typing derivation for the term $(\lambda x.e)\,v$ before reduction:
\begin{equation*}
\inference
  { \inference
  { \inference
      {\coctx{\Gamma, x:\tau_1}{\cclrd{r}\;\czip\;\cclrd{t}} \vdash e : \tau_2}
      {\coctx{\Gamma}{\cclrd{r}} \vdash \lambda x.e : \tau_1 \xrightarrow{\cclrd{t}} \tau_2 } &
   \coctx{\Gamma}{\cunit} \vdash v : \tau_1 }
  {\coctx{\Gamma}{\cclrd{r} \;\cpar\; (\cunit \,\cseq\, \cclrd{t})} \vdash (\lambda x.e)\,v : \tau_2} }
  {\coctx{\Gamma}{\cclrd{r} \;\cpar\; \cclrd{t}} \vdash (\lambda x.e)\,v : \tau_2}
\end{equation*}
The last step simplifies the coeffect annotation using the fact that $\cunit$ is a unit of $\cseq$.
From Lemma~\ref{thm:flat-subst-cbv} $\subst{e}{x}{v}$ has the same coeffect annotation as $e$.
As $\cclrd{r}\,\czip\,\cclrd{t}\;\cleq\;\cclrd{r}\,\cpar\,\cclrd{t}$, we can apply sub-coeffecting:
%
\begin{equation*}
\tyrule{sub}
 { \coctx{\Gamma}{\cclrd{r}\;\czip\;\cclrd{t}} \vdash \subst{e}{x}{v} : \tau_2 }
 { \coctx{\Gamma}{\cclrd{r}\;\cpar\;\cclrd{t}} \vdash \subst{e}{x}{v} : \tau_2 } 
\end{equation*}
%
Thus, the reduction preserves type and coeffect annotation (although this may not be the
\emph{only} typing of the original term).
\end{proof}

@@

In the call-by-name reduction of $(\lambda x.e_1)~e_2$, the expression $e_2$ is substituted for all 
occurrences of the variable $v$ in an expression $e_1$. As discussed in 
Section~\ref{sec:flat-syntax-props}, the call-by-name strategy does not \emph{in general} preserve 
the type of a program, but it does preserve the typing in some interesting cases. The typing is
preserved for different reasons in two of our systems, so we briefly review the concrete examples.

\paragraph{Data-flow.} 
The type preservation property does not hold for data-flow. This case is similar to the example
shown earlier with effectful computations. As a minimal example, consider the substitution of 
$\kvd{prev}~z$ for a variable $y$ in a function $\lambda x.y$:
%
\begin{equation*}
\begin{array}{rclcl}
 \coctx{y\!:\!\tau_1,z\!:\!\tau_1}{0} &\narrow{\vdash}& \lambda x.y : \tau_1 \xrightarrow{0} \tau_2 & & (\textnormal{before}) \\
 \coctx{z\!:\!\tau_1}{1} &\narrow{\vdash}& \lambda x.\kvd{prev}~z : \tau_1 \xrightarrow{1} \tau_2 & & (\textnormal{after}) \\
\end{array}
\end{equation*}
%
After the substitution, the coeffect of the body is $1$. The rule for lambda abstraction requires
that $1=\mathit{min}(r,s)$ and so the least solution is to set both $r,s$ to $1$. The substitution
this affects the coeffects attached both to the function type and the overall context.

Semantically, the coeffect over-approximates the actual requirements -- the code does not access
previous value of the argument $x$. This cannot be caputred by a flat coeffect system, but can
be captured using the structural system discussed in Chapter~\ref{ch:structural}.

\paragraph{Implicit parameters.} In data-flow, there is no typing for the resulting expression that
preserves the type of the function. However, this is not the case for all systems. Consider substituting
an implicit parameter access $\ident{?p}$ for a variable $y$:
%
\begin{equation*}
\begin{array}{rclcl}
 \coctx{y\!:\!\tau_1}{\emptyset} &\narrow{\vdash}& \lambda x.y : \tau_1 \xrightarrow{\emptyset} \tau_2 & & (\textnormal{before}) \\
 \coctx{\emptyset}{ \{\ident{?p}\} } &\narrow{\vdash}& \lambda x.\ident{?p} : \tau_1 \xrightarrow{\emptyset} \tau_2 & & (\textnormal{after}) \\
\end{array}
\end{equation*}
%
The above shows one possible typing of the body -- one that does not change the coeffects of the 
function type and attaches all additional coeffects (implicit parameters) to the context. In case
of implicit parameters (and, more generally, systems with set-like annotations) this is always 
possible.

\paragraph{Liveness.} In liveness, the type preservation also holds, but for a different reason. Consider
substituting any expression $e$ of type $\tau_1$ with coeffects $\cclrd{r}$ for a variable $y$:
%
\begin{equation*}
\begin{array}{rclcl}
 \coctx{y\!:\!\tau_1}{\ident{L}} &\narrow{\vdash}& \lambda x.y : \tau_1 \xrightarrow{\ident{L}} \tau_2 & & (\textnormal{before}) \\
 \coctx{\emptyset}{\ident{L}} &\narrow{\vdash}& \lambda x.e : \tau_1 \xrightarrow{\ident{L}} \tau_2 & & (\textnormal{after}) \\
\end{array}
\end{equation*}
%
In the original expression, both the overall context and the function type are annotated with \ident{L}, 
because the body contains a variable access. An expression $e$ can always be treated as being annotated
with $\ident{L}$ (because $\ident{L}$ is the top element of the lattice) and so substitution does not
change any coeffects.

\paragraph{Reduction theorem.}
The above examples (implicit parameters and liveness) demonstrate two particular kinds of coeffect 
algebra for which typing preservation holds. Proving the type preservation separately provides 
more insight into how the systems work and so we choose to consider separately.

\begin{definition}
We call a flat coeffect algebra \emph{top-pointed} if $\cunit$ is the greatest (top) coeffect scalar
($\forall r \in \C \;.\; r \,\cleq\, \cunit $) and \emph{bottom-pointed} if it is the smallest (bottom) 
element ($\forall r \in \C \;.\; r \,\cgeq\, \cunit $).
\end{definition}

\noindent
Liveness is an example of top-pointed coeffects as variables are annotated with 
$\ident{L}$ and $\ident{D} \leq \ident{L}$, while implicit parameters and data-flow are examples
of bottom-pointed coeffects. For top-pointed flat coeffects, the substitution lemma holds without 
additional requirements:

\begin{lemma}[Top-pointed substitution]
\label{thm:cbn-substitution-top}
In a top-pointed flat coeffect calculus with an aglebra $(\C, \cseq, \cpar, \czip, \cunit, \czero, \cleq)$,
substituting an expression $e_s$ with arbitrary coeffects $\cclrd{s}$ for a variable $x$ in $e_r$ does
not change the coeffects of $e_r$:
%
\begin{equation*}
\begin{array}{l}
 \coctx{\Gamma}{\cclrd{s}} \vdash e_s : \tau_s \;\; \wedge \;\; 
   \coctx{\Gamma_1,  x : \tau_s, \Gamma_2}{ \cclrd{r}  } \vdash e_r : \tau_r\\
 \quad \Rightarrow \;\; \coctx{\Gamma_1,\Gamma,\Gamma_2}{ \cclrd{r} } \vdash \subst{e_r}{x}{e_s} : \tau_r
\end{array}
\end{equation*}
\end{lemma}

\begin{proof}
Using sub-coeffecting ($\cclrd{s} \,\cleq\, \cunit$) and a variation of Lemma~\ref{thm:flat-subst-cbv}.
\end{proof}
%
\noindent
As variables are annotated with the top element $\cunit$, we can substitute the term $e_s$ 
for any variable and use sub-coeffecting to get the original typing (because
$\cclrd{s} \,\cleq\, \cunit$).

In a bottom pointed coeffect system, substituting $e$ for $x$ increases the context 
requirements. However, if the system satisfies the strong condition that $\czip = \cseq = \cpar$ 
then the context requirements arising from the substitution can be associated with the context
$\Gamma$, leaving the context requirements of a function value unchanged. As a result, substitution 
does not break soundness as in effect systems. The requirement $\czip = \cseq = \cpar$ holds for 
our implicit parameters example (all three operators are set union) and for other set-like 
coeffects. It allows the following substitution lemma:

\begin{lemma}[Bottom-pointed substitution]
\label{thm:cbn-substitution-bot}
In a bottom-pointed flat coeffect calculus with an algebra $(\C, \cseq, \cpar, \czip, \cunit, \czero, \cleq)$ 
where $\czip = \cseq = \cpar$ and the operation is also idempotent and commutative:
%
\begin{equation*}
\begin{array}{l}
\coctx{\Gamma}{\cclrd{s}} \vdash e_s : \tau_s \;\; \wedge \;\; 
\coctx{\Gamma_1,  x : \tau_s, \Gamma_2}{ \cclrd{r}  } \vdash e_r : \tau_r\\
\quad \Rightarrow \;\; \coctx{\Gamma_1,\Gamma,\Gamma_2}{ \cclrd{r} \,\cseq\, \cclrd{s} } \vdash \subst{e_r}{x}{e_s} : \tau_r
\end{array}
\end{equation*}

\end{lemma}
\begin{proof}
By induction over $\vdash$, using the idempotent, commutative monoid structure to keep $\cclrd{s}$ with
the free-variable context. See Appendix~?.
\end{proof}

\noindent
The flat system discussed here is \emph{flexible enough} to let us always re-associate new context 
requirements (arising from the substitution) with the free-variable context. In contrast, the 
structural system discussed in Chapter~\ref{ch:structural} is \emph{precise enough} to keep the 
coeffects associated with individual variables -- thus preserving typing in a complementary way.

The two substitution lemmas show that the call-by-name evaluation strategy can be used for
certain coeffect calculi, including liveness and implicit parameters. Assuming
$\reduce_{\ident{cbn}}$ is the standard call-by-name reduction, the following subject reduction
theorem holds:

\begin{theorem}[Call-by-name reduction]
In a coeffect system that satisfies the conditions for Lemma~\ref{thm:cbn-substitution-top} or
Lemma~\ref{thm:cbn-substitution-bot}, if $\coctx{\Gamma}{\cclrd{r}} \vdash e : \tau$ 
and $e \rightarrow_{\ident{cbn}} e'$ then $\coctx{\Gamma}{\cclrd{r}} \vdash e' : \tau$.
\end{theorem}
\begin{proof}

For top-pointed coeffect algebra (using Lemma~\ref{thm:cbn-substitution-top}), the proof is similar
to the one in Theorem~\ref{thm:cbv-reduction}, using the facts that $\cclrd{s} \cleq \cunit$
and $r \,\czip\, t \;=\; r \,\cpar\, t$.
For bottom-pointed coeffect algebra, consider the typing derivation for the term
$(\lambda x.e)\,v$ before reduction:
%
\begin{equation*}
\inference
  { \inference
      {\coctx{\Gamma, x:\tau_s}{\cclrd{r}} \vdash e_r : \tau_r}
      {\coctx{\Gamma}{\cclrd{r}} \vdash \lambda x.e_r : \tau_s \xrightarrow{\cclrd{r}} \tau_r } &
   \coctx{\Gamma}{\cclrd{s}} \vdash e_s : \tau_s }
  {\coctx{\Gamma}{\cclrd{r} \;\cpar\; (\cclrd{s} \,\cseq\, \cclrd{r})} \vdash (\lambda x.e_r)\,e_s : \tau_r }
\end{equation*}
%
The derivation uses the idempotence of $\czip$ in the first step, followed by the 
(\emph{app}) rule. The type of the term after substitution, using Lemma~\ref{thm:cbn-substitution-bot} is:
%
\begin{equation*}
\inference
  { \coctx{\Gamma, x:\tau_s}{\cclrd{r}} \vdash e_r : \tau_r & 
    \coctx{\Gamma}{\cclrd{s}} \vdash e_s : \tau_s }
  { \coctx{\Gamma, x:\tau_r}{\cclrd{r}\;\cseq\;\cclrd{s}} \vdash \subst{e_r}{x}{e_s} : \tau_s }
\end{equation*}
%
From the assumptions of Lemma~\ref{thm:cbn-substitution-bot}, we know that $\cseq\,=\,\cpar$
and the operation is idempotent, so trivially:
$\cclrd{r}\;\cseq\;\cclrd{s} = \cclrd{r} \;\cpar\; (\cclrd{s} \,\cseq\, \cclrd{r})$
\end{proof}

\paragraph{Expansion theorem.}
The $\eta$-expansion (local completeness) is similar to $\beta$-reduction (local soundness) in that
it holds for some flat coeffect systems, but not for all. Out of the examples we discuss, it 
holds for implicit parameters, but does not hold for liveness and data-flow.

Recall that $\eta$-expansion turns $e$ into $\lambxa x.e~x$. In the case of liveness, the 
expression $e$ may require no variables (both immediate and latent coeffects are marked as
$\ident{D}$). However, the resulting expression $\lambda x.e~x$ accesses a variable, 
marking the context and function argument as live. In case of data-flow, the coeffects are
made larger by the lambda abstraction. We remedy this limitation in the next chapter.

However, the $\eta$-expansion preserves the type for implicit parameters and, more generally,
for any flat coeffect algebra where $\cpar \,=\, \czip$. Assuming $\rightarrow_\eta$ is the 
standard $\eta$-reduction:

\begin{theorem}[$\eta$-expansion]
\label{thm:cbn-substitution-bot}
In a bottom-pointed flat coeffect calculus with an algebra $(\C, \cseq, \cpar, \czip, \cunit, \czero, \cleq)$ 
where $\czip = \cpar$, if $\coctx{\Gamma}{\cclrd{r}} \vdash e : \tau_1 \xrightarrow{\cclrd{s}} \tau_2$ 
and $e \rightarrow_\eta e'$ then $\coctx{\Gamma}{\cclrd{r}} \vdash e' : \tau_1 \xrightarrow{\cclrd{s}} \tau_2 $.
\end{theorem}
\begin{proof}
The following derivation shows that $\lambda x.e~x$ has the same type as $e$:
\begin{equation*}
\inference
  { \inference
    { \coctx{\Gamma}{\cclrd{r}} \vdash e : \tau_1 \xrightarrow{\cclrd{s}} \tau_2 &
      \coctx{x:\tau_1}{\cunit} \vdash x:\tau_1 }
    { \coctx{\Gamma, x:\tau_1}{\cclrd{r} \;\cpar\; (\cunit \,\cseq\, \cclrd{s}) } \vdash e~x : \tau_2 } }
  { \inference
    { \inference
      { \coctx{\Gamma, x:\tau_1}{\cclrd{r} \;\cpar\; \cclrd{s} } \vdash e~x : \tau_2 }
      { \coctx{\Gamma, x:\tau_1}{\cclrd{r} \;\czip\; \cclrd{s} } \vdash e~x : \tau_2 } }
    { \coctx{\Gamma}{ \cclrd{r} } \vdash \lambda x.e~x : \tau_1 \xrightarrow{\cclrd{s}} \tau_2 } }
\end{equation*}
\end{proof}

\noindent
The derivation starts with the expression $e$ and derives the type for $\lambda x.e~x$. The
application yields context requirements $\cclrd{r}\,\cpar\,\cclrd{s}$. In order to recover the
original typing, this must be equal to $\cclrd{r}\,\czip\,\cclrd{s}$. Note that the derivation
is showing just one possible typing -- the expression $\lambda x.e~x$ has other types -- but
this is sufficient for showing type preservation.

In summary, flat coeffect calculi do not \emph{in general} permit call-by-name evaluation, but
there are several cases where call-by-name evaluation can be used. Among the examples we discuss,
these include liveness and implicit parameters. Moreover, for implicit parameters (and more
generally, any set-like flat coeffect algebra), the $\eta$-expansion holds as well, giving us
both local soundness and local completeness as coined by Pfenning and Davies \cite{logic-modal-reconstruction}.

\subsection{Normal form?}
\label{sec:flat-syntax-norm}

@@

%
% (e2 + (e1 * e2)) + (e3 * f)
% ----(glet)------
% ----------(app)------------
%
% = e2 + e1*e2 + e3*f

% 
% (e2 + (e3 * f)) + ((e2 + (e3 * f)) * e1)
% ----(app)------    ----(app)-----
% ---------------(glet)-------------------
%
% = e2 + e3*f + e2*f + e1*e3*f

% Data-flow: (e1 + e2) max (e3 + f) = e1 + (e2 max (e3 + f))    XXXXX
% Liveness:  e2 ^ (e3 v f) = e2 ^ (e3 v f)
% Resources: OK

@@

% e3 + ((e2 + (e1 * e2)) * e3)
%       ------(glet)---
% ---------(glet)-------------
%
% = e3 + e2*e3 + e1*e2*e3

% (e2 + (e3 * e2)) + (e1 * (e2 + (e3 * e2)))
% -----(glet)-----         -----(glet)-----
% ------------------(glet)------------------
%
% = e2 + e3*e2 + e1*e2 + e1*e2*e3

% Data-flow: e3 + (e2 + e1) = (e2 + e3) + e1
% Liveness:  e3 = e3
% Resources: OK
